Clustering in a hyperbolic model of complex networks.

### 2019-03-13 concentration_argument.tex
finished revision of sections 5 and 6 and Appendix D

In the proof of Proposition~\ref{prop:asymp} we used a result that allowed us to restrict integration over $y$ to the interval $[a(k)^-, a(k)^+]$, with
\[
	a(k)^\pm = 2\log\left(\frac{k \pm C \sqrt{k \log(k)}}{\xi} \vee 1\right).
\]
The reason for this was that the integrand included the function $\rho(y,k) = \Prob{\Po(\mu(y)) = k}$, where $\Po(\lambda)$ denotes a Poisson random variable with expectation $\lambda$ and Poisson random variables are well concentrated around their mean, i.e. around heights $y$ for which $\mu(y) \approx k$. Since $\mu(y) = \mu(\BallPo{y}) = \xi e^{y/2}$, this implies that integration with respect to $\rho(y,k)$ is concentrated around $y \approx 2\log(k/\xi)$.

In the remainder of this paper we will often encounter integrands involving the function $\Prob{\Po(\mu_n(y)) = k}$, for some $\mu_n(y)$. In these case we want to be able to restrict our integration around those heights $y$ for which $y \approx 2\log(k_n/\xi)$. We will refer to such results as \emph{concentration of heights arguments}. In this section we establish such results. We start with a concentration of heights lemma for the infinite model $\Ginf$ (Lemma~\ref{lem:concentration_argument}) and explain in Remark~\ref{rmk:concentration_argument} how such a result will be used throughout the paper. We then establish similar results for the Poissonized KPKVB and finite box model based on a generalization of Lemma~\ref{lem:concentration_argument} obtained in Section~\ref{ssec:general_concentration_lemma}. To apply this lemma to $\GPo$ and $\Gbox$ we have to analyze the expected number of nodes in a typical neighborhood in these models, which is done in Sections~\ref{ssec:average_degree_P_n} and~\ref{ssec:average_degree_HP_n}.

@@

The next lemma states that for a large class of functions $h(y)$ and $k_n \to \infty$, to compute the integral 
\[
	\int_{0}^\infty \rho(y,k_n) h(y) e^{-\alpha y} \dd y
\]
%\TM{ This part is supposed to be for the infinite model, yet you integrate over the finite box $\Rcal_n$. Is this really meant?
%I would have expected we just integrate wrt.~$y$ and replace $f_{\alpha,\nu}(x,y)$ by $\alpha e^{-\alpha y}$, the exponential density.}\PvdH{I understand this confusion and agree that your suggestion makes more sense here. However, I choose this setup to minimize notations later. I am for changing the setup to what you are suggesting but I will have to spend some time first on checking how that will effect the definitions in later sections.}
it is enough to consider integration over a small interval on which $e^{y/2} \approx k_n$, instead of $\R_+$.

@@

Note that we can tune the error in~\eqref{eq:error_bound_int_rho_not_K} by selecting an appropriately large $C > 0$, i.e. by restrict the function $h(y)$ inside the integral to an appropriate interval around $2\log(k_n/\xi)$. This makes Lemma~\ref{lem:concentration_argument} very powerful. Below we list several important corollaries.

@@

\begin{proof}
By assumption, $g$ satisfies the conditions of for the second statement of Lemma~\ref{lem:concentration_argument}. Since in addition $g(2\log(k_n/\xi)) \sim h(2\log(k_n/\xi))$, the result follows.
\end{proof}

For any positive sequence $a_n \to \infty$ and $C > 0$ we define
\begin{equation}\label{eq:def_K_C_set}
	\Kcal_C(a_n) = \left\{y \in \R_+ : \frac{a_n - C \sqrt{a_n \log(a_n)}}{\xi} \vee 1 \le e^{\frac{y}{2}}
	\le \frac{a_n + C \sqrt{a_n \log(a_n)}}{\xi} \right\}.
\end{equation}
In addition we define
\begin{equation}\label{eq:def_K_C_n_set}
	\Kcal_{C,n}(a_n) := (-I_n, I_n] \times ((0,R_n] \cap \Kcal_{C}(a_n)),
\end{equation}
where $I_n := \frac{\pi}{2}e^{R_n/2}$. Recall that $\Rcal = (-I_n, I_n] \times [0,R]$. The following corollary allows us to bound integrals of function by considering there maximum of $\Kcal_{C}(k_n)$.

@@

@@

\begin{remark}[Proof of Proposition~\ref{prop:asymp} revisited]
Note that due to Proposition~\ref{prop:asymptotics_P}, the function $P(y)$ from Section~\ref{sec:asymptotics_average_clustering_ast_P} satisfies all the necessary conditions in Corollary~\ref{cor:concentration_of_heights_asymptotics}. Hence, Proposition~\ref{prop:asymp} directly follows from Proposition~\ref{prop:asymptotics_P} and a concentration of heights argument (second statement of Lemma~\ref{lem:concentration_argument}).
\end{remark}

Although powerful, the current versions of the concentration of heights arguments are only valid for the function $\rho(y,k_n) := \Prob{\Po\left(\Mu{\BallPo{y}}\right) = k_n}$, which uses the neighborhoods in the infinite model $\Ginf$. Since we will also be working in the Poissonized KPKVB model $\GPo$ and the finite box model $\Gbox$, we would like to use concentration of heights arguments when the integrand contains either the function $\rho_{\Po}(y,k_n) := \Prob{\Po\left(\Mu{\BallHyp{y}}\right) = k_n}$ or $\rho_{\text{box}}(y,k_n) := \Prob{\Po\left(\Mu{\BallPon{y}}\right) = k_n}$. The next proposition establishes this result and we will spend the rest of this section on its proof.

\begin{proposition}\label{prop:concentration_argument_all}
The conclusions from Lemmas~\ref{lem:concentration_argument} and Corollaries~\ref{cor:concentration_of_heights_asymptotics}, \ref{cor:concentration_heights_bounds_n} and~\ref{cor:concentration_heights_asymptotics_n} still hold if we replace $\rho(y,k_n)$ with either $\rho_{\Po}(y,k_n)$ or $\rho_{\text{box}}(y,k_n)$.
\end{proposition}

@@

\begin{lemma}\label{lem:concentration_argument_rho_approximation}
Let $\alpha > \frac{1}{2}, \nu > 0$ and $k_n \to \infty$ such that $k_n = \smallO{n}$, $\ell_n = (1 + \epsilon_n)k_n$, with $\epsilon_n \to 0$ and define
\[
	\lambda_n^\pm = (\ell_n \pm C \sqrt{\ell_n \log(\ell_n)}) \wedge \xi, \quad \text{and} \quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi}\right).
\] 
In addition, define $\hat{\rho}_n(y,k) = \Prob{\Po(\hat{\mu}_n(y)) = k}$, where $\hat{\mu}_n(y)$ satisfies, 
\[
	\hat{\mu}_n(y) = (1 + \phi_n(y))\mu(y),
\]
where $\phi_n(y)$ is a continuously differentiable function such that
\begin{enumerate}[\upshape i)]
\item For some $0 < \varepsilon < 1$, $\displaystyle \lim_{n \to \infty} \sup_{0 \le y \le (1 - \varepsilon)R_n} |\phi_n(y)| = 0$,
\item $\displaystyle \lim_{n \to \infty}  \sup_{y \in \Kcal_{C}(\ell_n)} |\phi_n^\prime(y)| = 0.$
\end{enumerate}
Then the following holds.
\begin{enumerate}
\item For any continuous function $h : \R_+ \rightarrow  \R$, such that $h(y) = \bigO{e^{\beta y}}$ as $y \to \infty$ for some $\beta < \alpha$, 
\begin{equation}\label{eq:error_bound_int_rho_approx_not_K}
	\int_{\R_+ \setminus (a_n^-, a_n^+)} \hat{\rho}_n(y,k_n) h(y) \alpha e^{-\alpha y} \dd y
	= \bigO{k_n^{-(1+C^2)/2}},
\end{equation}
as $n \to \infty$.
\item If in addition, $h(a_n) \sim h(b_n)$ whenever $a_n \sim b_n$, as $n \to \infty$, then,
\begin{equation}\label{eq:concentration_h_rho_approx}
	\int_0^\infty h(y) \hat{\rho}_n(y,k_n) \alpha e^{-\alpha y} \dd y \sim  
		2\alpha \xi^{2\alpha} h(2\log(k_n/\xi)) k_n^{-(2\alpha + 1)},
\end{equation}
as $n \to \infty$.
\end{enumerate}
\end{lemma}

@@

With this lemma, the proof of Proposition~\ref{prop:concentration_argument_all} follows if we can show that both $\rho_{\Po}(y,k_n)$ or $\rho_{\text{box}}(y,k_n)$ are of the form $\rho(y,k_n)(1 + \phi_n(y))$, where $\phi_n(y)$ satisfies the conditions i) and ii). We deal with each of these cases separately.

@@

The following lemma immediately implies that $\Mu{\BallPon{y}}$ satisfies the conditions of Lemma~\ref{lem:concentration_argument_rho_approximation} and hence that Proposition~\ref{prop:concentration_argument_all} holds for $\rho_{\text{box}}(y,k_n)$.

@@

\begin{proof}
First note that since we have identified the boundaries of $[-\frac{\pi}{2}e^{\frac{R_n}{2}}, \frac{\pi}{2}e^{\frac{R_n}{2}}]$ we can assume, without loss of generality, that $p = (0,y)$. We then have that the boundaries of $\BallPon{p}$ are given by the equations $x^\prime = \pm e^{\frac{y+y^\prime}{2}}$, which intersect the left and right boundaries of $[-\frac{\pi}{2}e^{\frac{R_n}{2}}, \frac{\pi}{2}e^{\frac{R_n}{2}}]$ at height
\[
	h(y) = R_n + 2 \log\left(\frac{\pi}{2}\right) - y.
\]
Therefore, if $y \le 2 \log(\pi/2)$ this intersection occurs above the height $R_n$ of the box $\Rcal_n$ while in the other case the full region of the box above $h(y)$ is connected to $p$.

@@

@@

We will now show that a concentration of heights argument also applies to the KPKVB model. Due to the hyperbolic distance formula, the computations are however more involved than for the finite box model. Recall that under the coupling between the hyperbolic random graph and the finite box model, for two points $p, p^\prime$ with $y + y^\prime < R_n$, $p^\prime \in \BallHyp{p}$ exactly when $|x-x^\prime|_{\pi e^{R_n/2}} \le \Phi(y,y^\prime)$, where we slightly abused notation to write $\Phi(y,y^\prime) = \Phi(R - y, R - y^\prime)$. In this setting, the coupling lemma (Lemma~\ref{lem:asymptotics_Omega_hyperbolic}) gives that  
\[
	e^{\frac{1}{2}(y+y^\prime)} - K e^{\frac{3}{2}(y+y^\prime) - R_n} \leq \Phi(y, y^\prime) 
		\leq  e^{\frac{1}{2}(y+y^\prime)} + K e^{\frac{3}{2}(y+y^\prime) - R_n},
\]
for some constant $K$. This result enables us to determine the measure of a ball around a given point $p=(0,y)$. Recall that the hyperbolic ball $\BallHyp{p}$ is a subset of $\Rcal_n$ and not of the hyperbolic disc $\Dcal_{R_n}$, i.e. the balls $\BallHyp{p}$ "live" in the finite box and not the hyperbolic disc. We start with the following preliminary result.

\begin{lemma}\label{lem:hyperbolic_ball_lower_part}
Let $\Phi(y,y^\prime) := \Phi(R - y,R - y^\prime)$ be defined as in~\eqref{eq:def_Omega_hyperbolic}. Then, for any $0 \le \delta < 1$
\[
	\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R_n} \Mu{\BallPo{y}}^{-1} \frac{2\nu \alpha}{\pi} \int_0^{(1-\delta)(R-y)} \Phi(y,y^\prime) e^{-\alpha y^\prime} = 1.
\]
\end{lemma}

@@

\begin{lemma}\label{lem:average_degree_hyperbolic}
For any $0 < \varepsilon < 1$
\[
	\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R_n} \frac{\Mu{\BallHyp{0,y}}}{\Mu{\BallPo{0,y}}} = 1.
\]
%Let $\eps \in (0, 1)$. Then for all $0 \le y \le (1 - \eps)R_n$
%\[
%	 1 - \phi_{\H,n}^{(1)}(y) - \phi_{\H,n}^{(2)}(y) \le \frac{\Mu{\BallHyp{0,y}}}{\Mu{\BallPo{0,y}}} 
%	 \le 1 - \phi_{\H,n}^{(1)}(y) + \phi_{\H,n}^{(2)}(y),
%\]
%where
%\[
%	\phi_{\H,n}^{(1)}(y) = \frac{2\alpha - 1 - 4\pi}{4\pi} e^{-(\alpha - \frac{1}{2})(R_n - y)} 
%		+ \left(\alpha - \frac{1}{2}\right)\pi e^{-(\alpha - \frac{1}{2})R_n - y/2},
%\]
%and
%\[
%	\phi_{\H,n}^{(2)}(y) = 
%	+ \begin{cases}
%			\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
%			&\mbox{if } 1/2 < \alpha < 3/2,\\
%			\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
%			\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
%			&\mbox{if } \alpha > 3/2,
%		\end{cases}
%\]
%with $K$ being the constant coming from the approximation of $\Phi$ in Lemma~\ref{lem:asymptotics_Omega_hyperbolic}.
%%\TM{ We want to say in the lemma where $K$ comes from. (not everyone reads surrounding text, or reads linearly, in a paper as long as this one.) }
\end{lemma}

\begin{proof}
We perform the computation of $\Mu{\BallHyp{0,y}}$ by splitting the integration with respect to the height $y^\prime$ into the cases $y^\prime > R_n - y$ and $y^\prime \le R_n - y$,
\[
	\Mu{\BallHyp{y}} 
	= \Mu{\BallHyp{y} \cap \Rcal_n ([0,R_n - y))} + \Mu{\BallHyp{y} \cap \Rcal_n ([R_n - y, R_n])}.
\]

For the first part we have that
\[
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} = \frac{2\nu \alpha}{\pi} \int_0^{R-y} \Phi(y,y^\prime) e^{-\alpha y^\prime}
	\dd y^\prime.
\]
Hence, by applying Lemma~\ref{lem:hyperbolic_ball_lower_part} with $\delta = 0$ we conclude that
\[
	\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R_n} \Mu{\BallPo{y}}^{-1}
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} = 1.
\]

For the second part we observer that $\BallHyp{(0,y)} \cap \Rcal_n ([R_n - y, R_n])= \Rcal_n([R_n-y,R_n])$. 
Thus, 
\begin{align*}
	&\hspace{-30pt}\Mu{\BallHyp{(0,y)} \cap \Rcal_n ([R_n - y, R_n])} \\
	&= \int_{R_n - y}^{R_n} \int_{I_n} f_{\alpha,\nu}(x^\prime, y^\prime) \dd x^\prime \dd y^\prime
		= \nu \alpha e^{R_n/2}\left(e^{-\alpha(R_n - y)} - e^{-\alpha R_n}\right)\\
	&= \Mu{\BallPo{0,y}} \frac{2\alpha - 1}{4\pi} \left( e^{-(\alpha - \frac{1}{2})(R_n - y)}
		- e^{-(\alpha - \frac{1}{2})R_n - y/2}\right), \numberthis \label{eq:mu_hyperbolic_ball_part_1}
\end{align*}
from which we conclude that
\[
	\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R_n} \Mu{\BallPo{y}}^{-1} 
	\Mu{\BallHyp{(0,y)} \cap \Rcal_n ([R_n - y, R_n])} = 0,
\]
which finishes the proof.
\end{proof}

>>>>

Here we show that if we integrate with respect to a function $\hat{\rho}_n(y,k_n) = \Prob{\Po(\mu_n(y)) = k_n}$ then we may restrict integration with respect to the \emph{height} $y$ to an interval on which $\mu_n(y) = \bigT{k_n}$. We will refer to such a result as a \emph{concentration of heights} result. In addition, if $\mu_n(y)$ is equivalent to $\mu(y)$ on this interval, then we may replace $\hat{\rho}_n(y,k_n)$ in the integral with $\rho(y,k_n) = \Prob{\Po(\mu(y)) = k_n}$ (the degree distribution of the typical point in $\Ginf$).

We start with a concentration of heights result for the infinite model $\Ginf$ (Lemma~\ref{lem:concentration_argument}) and explain in Remark~\ref{rmk:concentration_argument} how such a result will be used throughout the paper. We then present a generalization of this result (Lemma~\ref{lem:concentration_argument}) and use this to establish concentration of heights results for the Poissonized KPKVB $\GPo$ and finite box model $\Gbox$.

Finally we provide a general result that allow to substitute $\hat{\rho}_n(y,k_n)$ in the integrand with $\rho(y,k_n)$ and show that this holds in particular for the degree distributions in $\GPo$ and $\Gbox$, given by, respectively $\rho_{\Po}(y,k_n) := \Prob{\Po(\Mu{\BallHyp{y}}) = k_n}$ and $\rho_{\text{box}}(y,k_n) := \Prob{\Po(\Mu{\BallPon{y}}) = k_n}$.

@@

The next lemma states that for a large class of functions $h(y)$ and $k_n \to \infty$, to compute the integral 
\[
	\int_{0}^\infty \rho(y,k_n) h(y) e^{-\alpha y} \dd y
\]
%\TM{ This part is supposed to be for the infinite model, yet you integrate over the finite box $\Rcal$. Is this really meant?
%I would have expected we just integrate wrt.~$y$ and replace $f_{\alpha,\nu}(x,y)$ by $\alpha e^{-\alpha y}$, the exponential density.}\PvdH{I understand this confusion and agree that your suggestion makes more sense here. However, I choose this setup to minimize notations later. I am for changing the setup to what you are suggesting but I will have to spend some time first on checking how that will effect the definitions in later sections.}
it is enough to consider integration over a small interval on which $e^{y/2} \approx k_n$, instead of $\R_+$.

@@

Note that we can tune the error in~\eqref{eq:error_bound_int_rho_not_K} by selecting an appropriately large $C > 0$, i.e. by restricting the function $h(y)$ inside the integral to an appropriate interval around $2\log(k_n/\xi)$. This makes Lemma~\ref{lem:concentration_argument} very powerful. Below we list several important corollaries.

@@

\begin{proof}
By assumption, $g$ satisfies the conditions of the second statement of Lemma~\ref{lem:concentration_argument}. Since in addition $g(2\log(k_n/\xi)) \sim h(2\log(k_n/\xi))$, the result follows.
\end{proof}

%For any positive sequence $a_n \to \infty$ and $C > 0$ we define
%\begin{equation}\label{eq:def_K_C_set}
%	\Kcal_C(a_n) = \left\{y \in \R_+ : \frac{a_n - C \sqrt{a_n \log(a_n)}}{\xi} \vee 1 \le e^{\frac{y}{2}}
%	\le \frac{a_n + C \sqrt{a_n \log(a_n)}}{\xi} \right\}.
%\end{equation}
%In addition we define
%\begin{equation}\label{eq:def_K_C_n_set}
%	\Kcal_{C,n}(a_n) := (-I_n, I_n] \times ((0,R_n] \cap \Kcal_{C}(a_n)),
%\end{equation}
%where $I_n := \frac{\pi}{2}e^{R_n/2}$. Recall that $\Rcal = (-I_n, I_n] \times [0,R]$. 
The following corollary allows us to bound integrals of function by considering there maximum of $\Kcal_{C}(k_n)$.

@@

For functions $h_n(y) = k_n^s h(y)$ we obtain an asymptotic equivalent expression for the associated integral.

@@

\begin{remark}[Proof of Proposition~\ref{prop:asymp} revisited]
Note that due to Proposition~\ref{prop:asymptotics_P}, the function $P(y)$ from Section~\ref{sec:asymptotics_average_clustering_ast_P} satisfies all the necessary conditions in Corollary~\ref{cor:concentration_of_heights_asymptotics} with
\[
	h(y) := \begin{cases}
		e^{-\frac{y}{2}(4\alpha - 2)} c_\alpha \xi^{4\alpha - 2} &\mbox{if } \frac{1}{2} < \alpha < \frac{3}{4},\\
		\frac{y}{2} e^{-\frac{y}{2}} &\mbox{if } \alpha = \frac{3}{4},\\
		e^{-\frac{y}{2}} \frac{\alpha - \frac{1}{2}}{\alpha - \frac{3}{4}} &\mbox{if } \alpha > \frac{3}{4}.
	\end{cases}
\] 
Hence, Proposition~\ref{prop:asymp} directly follows from Proposition~\ref{prop:asymptotics_P} and a concentration of heights argument (Corollary~\ref{cor:concentration_of_heights_asymptotics}).
\end{remark}

%
%
%The next proposition establishes this result and we will spend the rest of this section on its proof.
%
%\begin{proposition}\label{prop:concentration_argument_all}
%The conclusions from Lemmas~\ref{lem:concentration_argument} and Corollaries~\ref{cor:concentration_of_heights_asymptotics}, \ref{cor:concentration_heights_bounds_n} and~\ref{cor:concentration_heights_asymptotics_n} still hold if we replace $\rho(y,k_n)$ with either $\rho_{\Po}(y,k_n)$ or $\rho_{\text{box}}(y,k_n)$.
%\end{proposition}

@@

Although powerful, the current versions of the concentration of heights arguments are only valid for the function $\rho(y,k_n) := \Prob{\Po\left(\Mu{\BallPo{y}}\right) = k_n}$, which uses the neighborhoods in the infinite model $\Ginf$. Since we will also be working in the Poissonized KPKVB model $\GPo$ and the finite box model $\Gbox$, we would like to use concentration of heights arguments when the integrand contains either the function $\rho_{\Po}(y,k_n)$ or $\rho_{\text{box}}(y,k_n)$. To establish these result we first prove a more general version of Lemma~\ref{lem:concentration_argument}.

\begin{lemma}\label{lem:concentration_argument_rho_approximation}
Let $\alpha > \frac{1}{2}, \nu > 0$ and $k_n \to \infty$ such that $k_n = \smallO{n}$, $\ell_n = (1 + \epsilon_n)k_n$, with $\epsilon_n \to 0$ and define
\[
	\lambda_n^\pm = (\ell_n \pm C \sqrt{\ell_n \log(\ell_n)}) \wedge \xi, \quad \text{and} \quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi}\right).
\] 
In addition, let $\hat{\rho}_n(y,k) = \Prob{\Po(\hat{\mu}_n(y)) = k}$, with $\hat{\mu}_n(y) = (1 + \phi_n(y))\mu(y)$ and 
where $\phi_n(y)$ is a differentiable function satisfying:
\begin{enumerate}[\upshape i)]
\item For some $0 < \varepsilon < 1$, $\displaystyle \lim_{n \to \infty} \sup_{0 \le y \le (1 - \varepsilon)R_n} |\phi_n(y)| = 0$,
\item $\displaystyle \lim_{n \to \infty}  \sup_{y \in \Kcal_{C}(\ell_n)} |\phi_n^\prime(y)| = 0.$
\end{enumerate}
Then the following holds:
\begin{enumerate}
\item For any continuous function $h : \R_+ \rightarrow  \R$, such that $h(y) = \bigO{e^{\beta y}}$ as $y \to \infty$ for some $\beta < \alpha$, 
\begin{equation}\label{eq:error_bound_int_rho_approx_not_K}
	\int_{\R_+ \setminus (a_n^-, a_n^+)} \hat{\rho}_n(y,k_n) h(y) \alpha e^{-\alpha y} \dd y
	= \bigO{k_n^{-(1+C^2)/2}},
\end{equation}
as $n \to \infty$.
\item If in addition, $h(a_n) \sim h(b_n)$ whenever $a_n \sim b_n$, as $n \to \infty$, then,
\begin{equation}\label{eq:concentration_h_rho_approx}
	\int_0^\infty h(y) \hat{\rho}_n(y,k_n) \alpha e^{-\alpha y} \dd y \sim  
		\int_0^\infty h(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y,
\end{equation}
as $n \to \infty$.
\end{enumerate}
\end{lemma}

@@

To apply this lemma to $\rho_{\Po}(y,k_n)$ or $\rho_{\text{box}}(y,k_n)$ we need to show that both these functions are of the form $\rho(y,k_n)(1 + \phi_n(y))$, where $\phi_n(y)$ satisfies the conditions i) and ii) in Lemma~\ref{lem:concentration_argument_rho_approximation}. We will do this in the next two sections and establish that the results from Corollaries~\ref{cor:concentration_of_heights_asymptotics}, \ref{cor:concentration_heights_bounds_n}
and~\ref{cor:concentration_heights_asymptotics_n} hold when we replace $\rho(y,k_n)$ with either $\rho_{\text{box}}(y,k_n)$
or $\rho_{\Po}(y,k_n)$.

\begin{remark}[Generalized concentration of heights arguments]
Since the results from Corollaries~\ref{cor:concentration_of_heights_asymptotics}, \ref{cor:concentration_heights_bounds_n} and~\ref{cor:concentration_heights_asymptotics_n} hold for any of the three functions $\rho(y,k_n)$, $\rho_{\text{box}}(y,k_n)$ and $\rho_{\Po}(y,k_n)$, we will will refer only to one of these three when using a concentration of heights argument of for any of the three models $\GPo$, $\Gbox$ and $\Ginf$. 
\end{remark}

@@

The following lemma immediately implies that $\Mu{\BallPon{y}}$ satisfies the conditions of Lemma~\ref{lem:concentration_argument_rho_approximation}.

@@

\begin{proof}
First note that since we have identified the boundaries of $[-\frac{\pi}{2}e^{\frac{R_n}{2}}, \frac{\pi}{2}e^{\frac{R_n}{2}}]$ we can assume, without loss of generality, that $p = (0,y)$. We then have that the boundaries of $\BallPon{p}$ are given by the equations $x^\prime = \pm e^{\frac{y+y^\prime}{2}}$, which intersect the left and right boundaries of $[-\frac{\pi}{2}e^{\frac{R_n}{2}}, \frac{\pi}{2}e^{\frac{R_n}{2}}]$ at height
\[
	h(y) = R_n + 2 \log\left(\frac{\pi}{2}\right) - y.
\]
Therefore, if $y \le 2 \log(\pi/2)$ this intersection occurs above the height $R_n$ of the box $\Rcal$ while in the other case the full region of the box above $h(y)$ is connected to $p$.

@@

The following is now immediate.

\begin{corollary}\label{cor:concentration_of_heights_Gbox}
The statement in Corollaries~\ref{cor:concentration_of_heights_asymptotics}, \ref{cor:concentration_heights_bounds_n}
and~\ref{cor:concentration_heights_asymptotics_n} hold when we replace $\rho(y,k_n)$ with $\rho_{\text{box}}(y,k_n)$.
\end{corollary}

@@

We will now show that a concentration of heights argument also applies to the KPKVB model. Due to the hyperbolic distance formula, the computations are however more involved than for the finite box model. Recall that under the coupling between the hyperbolic random graph and the finite box model, for two points $p, p^\prime$ with $y + y^\prime < R_n$, $p^\prime \in \BallHyp{p}$ exactly when $|x-x^\prime|_{\pi e^{R_n/2}} \le \Phi(y,y^\prime)$. In this setting, the coupling lemma (Lemma~\ref{lem:asymptotics_Omega_hyperbolic}) gives that  
\[
	e^{\frac{1}{2}(y+y^\prime)} - K e^{\frac{3}{2}(y+y^\prime) - R_n} \leq \Phi(y, y^\prime) 
		\leq  e^{\frac{1}{2}(y+y^\prime)} + K e^{\frac{3}{2}(y+y^\prime) - R_n},
\]
for some constant $K$. This result enables us to determine the measure of a ball around a given point $p=(0,y)$. Recall that the hyperbolic ball $\BallHyp{p}$ is a subset of $\Rcal$ and not of the hyperbolic disc $\Dcal_{R_n}$, i.e. the balls $\BallHyp{p}$ "live" in the finite box and not the hyperbolic disc. We start with the following preliminary result.

\begin{lemma}\label{lem:hyperbolic_ball_lower_part}
Let $\Phi(y,y^\prime)$ be defined as in~\eqref{eq:def_Omega_hyperbolic}. Then, for any $0 \le \delta < 1$
\[
	\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R_n} \Mu{\BallPo{y}}^{-1} \frac{2\nu \alpha}{\pi} \int_0^{(1-\delta)(R-y)} \Phi(y,y^\prime) e^{-\alpha y^\prime} = 1.
\]
\end{lemma}

@@

\begin{lemma}\label{lem:average_degree_hyperbolic}
For any $0 < \varepsilon < 1$
\[
	\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R_n} \frac{\Mu{\BallHyp{y}}}{\Mu{\BallPo{y}}} = 1.
\]
%Let $\eps \in (0, 1)$. Then for all $0 \le y \le (1 - \eps)R_n$
%\[
%	 1 - \phi_{\H,n}^{(1)}(y) - \phi_{\H,n}^{(2)}(y) \le \frac{\Mu{\BallHyp{0,y}}}{\Mu{\BallPo{0,y}}} 
%	 \le 1 - \phi_{\H,n}^{(1)}(y) + \phi_{\H,n}^{(2)}(y),
%\]
%where
%\[
%	\phi_{\H,n}^{(1)}(y) = \frac{2\alpha - 1 - 4\pi}{4\pi} e^{-(\alpha - \frac{1}{2})(R_n - y)} 
%		+ \left(\alpha - \frac{1}{2}\right)\pi e^{-(\alpha - \frac{1}{2})R_n - y/2},
%\]
%and
%\[
%	\phi_{\H,n}^{(2)}(y) = 
%	+ \begin{cases}
%			\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
%			&\mbox{if } 1/2 < \alpha < 3/2,\\
%			\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
%			\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
%			&\mbox{if } \alpha > 3/2,
%		\end{cases}
%\]
%with $K$ being the constant coming from the approximation of $\Phi$ in Lemma~\ref{lem:asymptotics_Omega_hyperbolic}.
%%\TM{ We want to say in the lemma where $K$ comes from. (not everyone reads surrounding text, or reads linearly, in a paper as long as this one.) }
\end{lemma}

\begin{proof}
We perform the computation of $\Mu{\BallHyp{0,y}}$ by splitting the integration with respect to the height $y^\prime$ into the cases $y^\prime > R_n - y$ and $y^\prime \le R_n - y$,
\[
	\Mu{\BallHyp{y}} 
	= \Mu{\BallHyp{y} \cap \Rcal ([0,R_n - y))} + \Mu{\BallHyp{y} \cap \Rcal ([R_n - y, R_n])}.
\]

For the first part we have that
\[
	\Mu{\BallHyp{(0,y)} \cap\Rcal[(0,R_n-y)]} = \frac{2\nu \alpha}{\pi} \int_0^{R-y} \Phi(y,y^\prime) e^{-\alpha y^\prime}
	\dd y^\prime.
\]
Hence, by applying Lemma~\ref{lem:hyperbolic_ball_lower_part} with $\delta = 0$ we conclude that
\[
	\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R_n} \Mu{\BallPo{y}}^{-1}
	\Mu{\BallHyp{(0,y)} \cap\Rcal[(0,R_n-y)]} = 1.
\]

For the second part we observer that $\BallHyp{(0,y)} \cap \Rcal ([R_n - y, R_n])= \Rcal([R_n-y,R_n])$. 
Thus, 
\begin{align*}
	&\hspace{-30pt}\Mu{\BallHyp{(0,y)} \cap \Rcal ([R_n - y, R_n])} \\
	&= \int_{R_n - y}^{R_n} \int_{I_n} f_{\alpha,\nu}(x^\prime, y^\prime) \dd x^\prime \dd y^\prime
		= \nu \alpha e^{R_n/2}\left(e^{-\alpha(R_n - y)} - e^{-\alpha R_n}\right)\\
	&= \Mu{\BallPo{0,y}} \frac{2\alpha - 1}{4\pi} \left( e^{-(\alpha - \frac{1}{2})(R_n - y)}
		- e^{-(\alpha - \frac{1}{2})R_n - y/2}\right), \numberthis \label{eq:mu_hyperbolic_ball_part_1}
\end{align*}
from which we conclude that
\[
	\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R_n} \Mu{\BallPo{y}}^{-1} 
	\Mu{\BallHyp{(0,y)} \cap \Rcal ([R_n - y, R_n])} = 0,
\]
which finishes the proof.
\end{proof}

A direct consequence of Lemma~\ref{lem:average_degree_hyperbolic} is that $\Mu{\BallHyp{y}} = \Mu{\BallPo{y}}(1 + \phi_n(y))$, where $\phi_n(y) := \Mu{\BallHyp{y}}/\Mu{\BallPo{y}} - 1$ satisfies condition i) in Lemma~\ref{lem:concentration_argument_rho_approximation}. To show that condition ii) is also satisfied we need to analyze
\[
	\phi_n^\prime(y) = 
	\Mu{\BallPo{y}}^{-1} \frac{\partial}{\partial y} \mu\left(\BallHyp{y}\right) -  \frac{1}{2} \frac{\Mu{\BallHyp{y}}}{\Mu{\BallPo{y}}},
\]
where we used that $\frac{\partial}{\partial y}\Mu{\BallPo{y}} = \frac{1}{2}\Mu{\BallPo{y}}$. Again, Lemma~\ref{lem:average_degree_hyperbolic} implies that 
\[
	\lim_{n \to \infty} \sup_{0 \le y \le (1-\varepsilon)R_n} \frac{1}{2} \frac{\Mu{\BallHyp{y}}}{\Mu{\BallPo{y}}}
	= \frac{1}{2}.
\]
The following lemma shows that the same holds for the first term from which we conclude that $\phi_n(y)$ satisfies condition ii) in Lemma~\ref{lem:concentration_argument_rho_approximation}.
